
\section{The Lusternik-Schnirelmann category}

As we have just developed a new way to test formality, it would be nice to test it out on 
some examples. Our gut feeling is that this criteria of having trivial induced product in 
cohomology is pretty strong. If we are hoping for results applicable to topological spaces 
this is especially troubling. Say we have a path-connected topological space $X$ — then it 
has zeroth cohomology $H^0(X;K)\cong K$. Hence the requirement to have trivial cup product 
reduces to requiring $H^i(X;K)=0$ for $i>0$, which is really limiting. One solution to this 
is looking at reduced cohomology.

\begin{definition}
    Let $X$ be a topological space and $C^*(X;K)$ its cochain complex (treated here as a 
    dg-algebra). We define its augmented cochain dg-algebra, denoted 
    $\widetilde{C}^*(X;K)$ by adding a copy of the ground field $K$ injectively farthest 
    to the left, i.e.
    $$\cdots \longrightarrow 0 \longrightarrow K \overset{\epsilon}\longrightarrow C^0(X)\longrightarrow C^1(X) \longrightarrow \cdots \longrightarrow C^n(X) \longrightarrow \cdots $$
\end{definition} 

The cohomology algebra of the augmented cochain complex is called the reduced cohomology 
algebra of $X$ and is denoted $\widetilde{H}^*(X;K)$. If the space $X$ is connected, 
then $\widetilde{H}^0(X;K)=0$, meaning that we have completely removed the problem 
described above.

Spaces with trivial cup product are not in abundance, but examples include the spheres 
and more generally the suspended spaces. There are ways to make sure that a space has a 
trivial cup product — one of these using the Lusternik-Schnirelmann category. 

\begin{definition}
    Let $X$ be a topological space. The Lusternik-Schnirelmann category of $X$, denoted 
    $\text{cat}_{LS}(X)$, is the least integer $n$ such that there is a covering of $X$ 
    by $n+1$ open subsets $U_i$ that are all contractible in $X$, i.e. their inclusion 
    into $X$ is null-homotopic.    
\end{definition}

This invariant was originally developed in \cite{lscat} as an invariant on manifolds to be 
a lower bound for the number of critical points any real valued function on it could have. 
It has since become a useful – but very difficult to calculate – invariant of topological 
spaces. 

So why do we introduce this invariant? Recall that the cup-length of a topological space 
$X$, denoted $\text{cl}(X)$ is the largest integer $n$ such that a chain 
$[x_1]\cup \cdots \cup [x_n]$ of cohomology classes with $\deg|x_i|\geq 1$ is non-zero. 
We have the following fundamental relation between the cup length and the 
Lusternik-Schnirelmann category of $X$ \cite{lscategorybook}.

\begin{lemma}
    Let $X$ be a topological space. Then the cup length of $X$ is a lower bound for its 
    Lusternik-Schnirelmann category, i.e. $\text{cl}(X)\leq \text{cat}_{LS}(X)$.    
\end{lemma}

Thus, choosing spaces with $\text{cat}_{LS}(X) = 1$ means we have trivial product on 
reduced cohomology. Examples of such spaces are again the suspended spaces, as they are 
the union of two cones. The last thing we need to know is whether spaces with 
Lusternik-Schnirelmann category $1$ admit any non-vanishing Massey  $n$-products. Luckily, 
by Rudyak we know that they dont.

\begin{theorem}[\cite{Rudyak}]
    Let $X$ be a topological space with $\text{cat}_{LS}(X)\leq 1$ and 
    $x_1, \ldots, x_n \in \widetilde{H}^\ast(X)$. If the Massey $n$-product 
    $\langle x_1, \ldots, x_n\rangle$ is defined then 
    $\langle x_1, \ldots, x_n\rangle = 0$.    
\end{theorem}

Call spaces with formal reduced cochain algebras reduced formal spaces. This means now
that spaces with Lusternik-Schnirelmann category $1$ are reduced formal by the above 
theorem together with the earlier result on formality. 

\begin{corollary}
    Let $X$ be a topological space with $\text{cat}_{LS}(X)=1$. Then 
    $\widetilde{C}^\ast(X)$ is a formal dg-algebra. 
\end{corollary}

We still have to deal with the degree $0$ cochains in order to call 
$X$ formal and not just reduced formal. We need that either its piece-wise linear De Rham 
algebra or its full – non-reduced – cochain algebra is formal. 

\begin{theorem}
    Any reduced formal topological space $X$ is formal.     
\end{theorem}

\begin{proof}
    Since $X$ is reduced formal we know that there is a span of dg-quasi-isomorphisms 
    $\widetilde{H}^\ast(X)\longleftarrow B \longrightarrow \widetilde{C}^\ast(X)$ for some 
    dg-algebra $B$:

    \begin{center}
        \begin{tikzpicture}
            \node (1) {$k$};
            \node (2) [right of=1] {$C^0(X)$};
            \node (3) [right of=2] {$C^1(X)$};
            \node (4) [right of=3] {$C^2(X)$};
            
            \node (5) [below of=1] {$B^{-1}$};
            \node (6) [right of=5] {$B^0$};
            \node (7) [right of=6] {$B^1$};
            \node (8) [right of=7] {$B^2$};
            
            \node (9) [below of=5] {$0$};
            \node (10) [right of=9] {$0$};
            \node (11) [right of=10] {$H^1(X)$};
            \node (12) [right of=11] {$H^2(X)$};
            
            \node (13) [left of=1] {$\cdots$};
            \node (14) [left of=5] {$\cdots$};
            \node (15) [left of=9] {$\cdots$};
            \node (16) [right of=4] {$\cdots$};
            \node (17) [right of=8] {$\cdots$};
            \node (18) [right of=12] {$\cdots$};
            
            \draw [-to] (13) to node {} (1);
            \draw [-to] (14) to node {} (5);
            \draw [-to] (15) to node {} (9);
            \draw [-to] (4) to node {} (16);
            \draw [-to] (8) to node {} (17);
            \draw [-to] (12) to node {} (18);
            
            \draw [-to] (1) to node {$\epsilon$} (2);
            \draw [-to] (2) to node {$d^0$} (3);
            \draw [-to] (3) to node {$d^1$} (4);
            
            \draw [-to] (5) to node {$d^{-1}_B$} (6);
            \draw [-to] (6) to node {$d^0_B$} (7);
            \draw [-to] (7) to node {$d^1_B$} (8);
            
            \draw [-to] (9) to node {} (10);
            \draw [-to] (10) to node {} (11);
            \draw [-to] (11) to node {$0$} (12);
            
            \draw [-to] (5) to node {$q^{-1}$} (1);
            \draw [-to] (5) to node {} (9);
            
            \draw [-to] (6) to node {$q^0$} (2);
            \draw [-to] (6) to node [swap]{$p^0$} (10);
            
            \draw [-to] (7) to node {$q^1$} (3);
            \draw [-to] (7) to node [swap]{$p^1$} (11);
            
            \draw [-to] (8) to node {$q^2$} (4);
            \draw [-to] (8) to node [swap]{$p^2$} (12);
        \end{tikzpicture}
        \end{center}

    By changing the diagram slightly at the left-most side, we get the following new 
    diagram:

    \begin{center}
        \begin{tikzpicture}
            \node (1) {$0$};
            \node (2) [right of=1] {$C^0(X)$};
            \node (3) [right of=2] {$C^1(X)$};
            \node (4) [right of=3] {$C^2(X)$};
            
            \node (5) [below of=1] {$0$};
            \node (6) [right of=5] {$B^0\oplus B^{-1}$};
            \node (7) [right of=6] {$B^1$};
            \node (8) [right of=7] {$B^2$};
            
            \node (9) [below of=5] {$0$};
            \node (10) [right of=9] {$k$};
            \node (11) [right of=10] {$H^1(X)$};
            \node (12) [right of=11] {$H^2(X)$};
            
            \node (13) [left of=1] {$\cdots$};
            \node (14) [left of=5] {$\cdots$};
            \node (15) [left of=9] {$\cdots$};
            \node (16) [right of=4] {$\cdots$};
            \node (17) [right of=8] {$\cdots$};
            \node (18) [right of=12] {$\cdots$};
            
            \draw [-to] (13) to node {} (1);
            \draw [-to] (14) to node {} (5);
            \draw [-to] (15) to node {} (9);
            \draw [-to] (4) to node {} (16);
            \draw [-to] (8) to node {} (17);
            \draw [-to] (12) to node {} (18);
            
            \draw [-to] (1) to node {} (2);
            \draw [-to] (2) to node {$d^0$} (3);
            \draw [-to] (3) to node {$d^1$} (4);
            
            \draw [-to] (5) to node {} (6);
            \draw [-to] (6) to node {$d^0_B$} (7);
            \draw [-to] (7) to node {$d^1_B$} (8);
            
            \draw [-to] (9) to node {} (10);
            \draw [-to] (10) to node {} (11);
            \draw [-to] (11) to node {$0$} (12);
            
            \draw [-to] (5) to node {} (1);
            \draw [-to] (5) to node {} (9);
            
            \draw [-to] (6) to node {$[q^0, 0]$} (2);
            \draw [-to] (6) to node [swap]{$[0, p^0]$} (10);
            
            \draw [-to] (7) to node {$q^1$} (3);
            \draw [-to] (7) to node [swap]{$p^1$} (11);
            
            \draw [-to] (8) to node {$q^2$} (4);
            \draw [-to] (8) to node [swap]{$p^2$} (12);
        \end{tikzpicture}
        \end{center}

    which gives a span of dg-quasi-isomorphisms 
    $H^\ast(X)\longleftarrow B'\longrightarrow C^\ast(X)$. Hence $X$ is formal. 
\end{proof}

We can then conclude with our result as a corollary.
\begin{corollary}
    Let $X$ be a space with $\text{cat}_{LS}(X)=1$. Then $X$ is formal.     
\end{corollary}

This result is certainly already known by specialists, but this seems to be a new method 
of proving it. Alternatively one can use that any space $X$ with $\text{cat}_{LS}(X)=1$ 
is a co-H-space \cite{hess}, and then that any co-H-space is a wedge spheres 
\cite{co-H-space}. Formality is preserved under the wedge product \cite{hess}, and since 
spheres are formal, we know that any co-H-space, and thus any space $X$ with 
$\text{cat}_{LS}(X)=1$ is a formal space. The special case of suspended spaces being 
formal was also proven in \cite{FHT}. 

\begin{question}
    Are there any formal spaces with trivial cup product and Lusternik-Schnirelmann 
    category greater than $1$?
\end{question}